---
title: MT4608 Sampling Theory
author: David Borchers
subtitle: Computer Practical 1
date: Septemeber 2020
output: 
  pdf_document:
    number_sections: true
fontsize: 12pt
classoption: a4paper
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Specify whether answers are shown 
#answer <- TRUE
answer <- FALSE
```

In this practical we are going to use the \texttt{R} package \texttt{survey} to reproduce some of the results we have calculated manually in class and in tutorials, and then to ...

# Getting started with the pacakge \texttt{sampling}

The \texttt{sampling} package should already be installed in RStudio Cloud, but if you are working on another machine, you need to install it from CRAN. 

We start by loading the \texttt{survey} package and the caribou dataset of Table 2.1 of the notes. The dataset is in the package \texttt{mt4608}, written specifically for this module^[This package is not as robust as packages on CRAN, so please be a bit patient/tolerant - things may go a little wrong in situations that I did not anticipate when writing it.].
```{r library, message=FALSE, warning=FALSE}
library(survey) # load survey package
library(mt4608) # load this module package
```

The caribou survey data are contained in the package \texttt{mt4608}, and can be accessed as follows:
```{r cariboudata, eval=FALSE}
data(caribou) # get the dataset
caribou # look at it
help(caribou) # get a description of it
```

Calculate the total area and total number of strips in the survey region:
```{r caribouN, eval=FALSE}
A = unique(caribou$area[caribou$stratum==1]) + 
  unique(caribou$area[caribou$stratum==2])
N = unique(caribou$N[caribou$stratum==1]) + 
  unique(caribou$N[caribou$stratum==2])
```

Estimate the mean number of animals per survey strip, together with a 95\% confidence interval. To do this with the \texttt{survey} package, you first need to associate a survey design with the survey data, using the command \texttt{svydesign}. Look at the help for  \texttt{svydesign} to understand more about the arguments; here  \texttt{ids=$\sim$1} tells it that there are no clusters (you have to specify \texttt{ids}), and \texttt{fpc} if the ``finite population correction'' factor that we call $f$ in the notes.
```{r cariboudesign, eval=FALSE}
n = dim(caribou)[1] # sample size
srs <- svydesign(ids=~1,data=caribou,fpc=rep(n/N,n))
```

Now using the survey data and the design, we can estimate the population mean and 95\% confidence interval very easily using the commands \texttt{svymean} and \texttt{confint}. To do this you have to tell \texttt{svymean} which column of the data frame is the $y$-variable (the response). In our case it is the \texttt{count}:
```{r caribouest, eval=FALSE}
ybar <- svymean(~count,srs)
ybar
confint(ybar)
```

## Exercise 1

\begin{enumerate}
\item By doing appropriate calculations with the data, decide whether the function \texttt{confint} assumed that the sample mean has a normal distribution or a t-distribution.

```{r ex1asoln, eval=FALSE, echo=FALSE}
s2 = sum((count-ybar)^2)/(n-1)
f = n/N
var.ybar = (1-f)*s2/n
var.ybar
tcrit = abs(qt(0.025,n-1))
ci.t = ybar[1] + c(-1,1)*tcrit*sqrt(var.ybar)
ci.t
ci.z = ybar[1] + c(-1,1)*1.96*sqrt(var.ybar)
ci.z
```

\item Estimate the total number of caribou in the survey region, together with a 95\% confidence interval

\end{enumerate}


```{r ex1bsoln, eval=FALSE, echo=FALSE}
# Assuming normality:
tot <- svytotal(~count,srs)
tot
round(confint(tot)) 

# Assuming t-distribution (better)
s2 = sum((count-ybar)^2)/(n-1)
f = n/N
var.tot = N^2*(1-f)*s2/n
var.tot
tcrit = abs(qt(0.025,n-1))
ci.t = tot[1] + c(-1,1)*tcrit*sqrt(var.tot) 
round(ci.t)
```

